Chatbots and messaging platforms in the classroom: an analysis from the teacher's perspective

### 2023-01-10 edubots_surveys.Rnw
Convert to footnote to keep URL

Several studies to evaluate chatbots in higher education have been conducted. For
instance, a comprehensive study was presented by Smutny and Schreiberova
\cite{Smutny2020} reviewing 47 educational chatbots implemented in Facebook
Messenger, concluding that most of these chatbots were very basic and that programmers require of more support to develop and 
offer tools for improving coaching methods and students learning. Pérez and collaborators
\cite{perez2020rediscovering} attempted to categorize educational chatbots attending to their purpose 
as service-oriented or teaching-oriented. Service-oriented chatbots
include those that provide service support, such as Ask Holly
\cite{meetholly} and Dina \cite{santoso2018dinus}, which respond to
students' questions about enrollment and registration. Ask L. U.
\cite{LancasterUniversity_2019} answers students' frequently asked questions
about schedules, grades, tutors, and societies. LISA \cite{dibitonto2018chatbot}
and Differ \cite{differ.chat} facilitate breaking the ice between new students
by introducing them to each other. Ranoliya et al. \cite{ranoliya2017chatbot}
proposed a generic chatbot for university students that is able to answer 
frequently asked questions. The University of Granada deployed
Elvira, a chatbot embedded in its main web page \cite{MOREO20129} to perform the same task. In addition to being able
to answer pre-established frequent questions, it did so from the website of
the University of Granada using an inset persona, voiced by a real person, who
lip-synced the answers. However, in the case of Elvira, the emphasis was on the
authenticity of the speech-mouth and face gestures synchronization rather than making updating
answered questions simpler or more interactive for the administrative staff, or more open to the rest of the university staff using actual messaging platforms. In
most cases, the embedded search engine was able to provide more up-to-date
and accurate answers than those provided by Elvira, which had to be updated
by hand (and not too often). It was eventually discontinued, and its technology was not adapted to other platforms, since it did not really provide a useful service. Its service was not tied to teaching anyway; it was more related to directory queries and university-wide administrative questions, so even if it is strictly an example of a chatbot in a university setting, it is not really a chatbot that can be adopted by teachers to affect the learning outcomes, which is the focus of this paper.

@@

% Describe the survey and population
We collected responses from 282 teachers from Spain and Spanish-speaking
countries for \textit{survey1} and 205 for \textit{survey2}. The two forms were
open for approximately the same amount of time, around two months. The sample
includes mostly university teachers, although about 32\% and 8\% are from
non-universitary tertiary education in the first and second survey respectively. Regarding
gender, 61\% of teachers were male and 32\% female, while approximately 5\%
chose not to disclose it. Finally, the responses are more or less equally
distributed according to the teaching experience, showing 24\% of responses from
teachers with 5 or less years of experience, 26\% for 6-15 years of experience,
30\% for 16-25, and 20\% for teachers with more than 25 years of experience.

We will try to find the answers to the four research questions next, by
analyzing the responses to the different survey questions.

@@

Regarding gender in Fig. \ref{figure:gender*postcovid_chisq}, the results show a
high positive contribution from female educators that did not use a messaging
app before the COVID-19 pandemic but do use it now; approximately 31\% of the
test results are explained by this cell. In contrast, male educators that
responded in the same way were less than expected (negative contribution that explains
17\% of the test results). Also, it is important to highlight the female
educators that continued to use these apps before and after the pandemic and
correspondingly the negative contribution of male educators to the same case
(less than expected by chance).

>>>>

Several studies to evaluate chatbots in higher education have been conducted. For
instance, a comprehensive study was presented by Smutny and Schreiberova
\cite{Smutny2020} reviewing 47 educational chatbots implemented in Facebook
Messenger, concluding that most of these chatbots were very basic and that programmers require of more support to develop and 
offer tools for improving coaching methods and students learning. Pérez and collaborators
\cite{perez2020rediscovering} attempted to categorize educational chatbots attending to their purpose 
as service-oriented or teaching-oriented. Service-oriented chatbots
include those that provide service support, such as Ask Holly
\cite{meetholly} and Dina \cite{santoso2018dinus}, which respond to
students' questions about enrollment and registration. Ask L. U.\footnote{"Lancaster University launch pioneering chatbot companion for students", {\tt https://www.lancaster.ac.uk/news/lancaster-university-launch-pioneering-chatbot-companion-for-students}.}
answers students' frequently asked questions
about schedules, grades, tutors, and societies. LISA \cite{dibitonto2018chatbot}
and Differ \cite{differ.chat} facilitate breaking the ice between new students
by introducing them to each other. Ranoliya et al. \cite{ranoliya2017chatbot}
proposed a generic chatbot for university students that is able to answer 
frequently asked questions. The University of Granada deployed
Elvira, a chatbot embedded in its main web page \cite{MOREO20129} to perform the same task. In addition to being able
to answer pre-established frequent questions, it did so from the website of
the University of Granada using an inset persona, voiced by a real person, who
lip-synced the answers. However, in the case of Elvira, the emphasis was on the
authenticity of the speech-mouth and face gestures synchronization rather than making updating
answered questions simpler or more interactive for the administrative staff, or more open to the rest of the university staff using actual messaging platforms. In
most cases, the embedded search engine was able to provide more up-to-date
and accurate answers than those provided by Elvira, which had to be updated
by hand (and not too often). It was eventually discontinued, and its technology was not adapted to other platforms, since it did not really provide a useful service. Its service was not tied to teaching anyway; it was more related to directory queries and university-wide administrative questions, so even if it is strictly an example of a chatbot in a university setting, it is not really a chatbot that can be adopted by teachers to affect the learning outcomes, which is the focus of this paper.

@@

We collected responses from 282 teachers from Spain and Spanish-speaking
countries for \textit{survey1} and 205 for \textit{survey2}. The two forms were
open for approximately the same amount of time, around two months. Most responders were university teachers, although about 32\% of them in \textit{survey1} and 8\% (in \textit{survey2} are from
non-universitary tertiary education teachers. With respect to
gender, 61\% of teachers were male and 32\% female, while approximately 5\%
chose not to disclose their gender. Finally, the responses are more or less equally
distributed according to the teacher's years of experience, showing 24\% of responses from
teachers with 5 or less years of experience, 26\% for 6-15 years of experience,
30\% for 16-25, and 20\% for teachers with more than 25 years of experience.

We will try to assess whether we found the answers to the four research questions next, by
analyzing the responses to the different survey questions.

@@

Regarding gender in Fig. \ref{figure:gender*postcovid_chisq}, the results show a
high positive contribution from female educators that did not use a messaging
app before the COVID-19 pandemic but do use it now; approximately 31\% of the
test results are explained by this cell. In contrast, male educators that
responded the same were less than expected (negative contribution that explains
17\% of the test results). Also, it is important to highlight the female
educators that continued to use these apps before and after the pandemic and
correspondingly the negative contribution of male educators to the same case
(less than expected by chance).
